---
title: "Urban Accidents in the City of Porto Alegre"
author: "Jean-Marc Vincent, Lucas Mello Schnorr"
date: "October 2017"
output:
  pdf_document:
    number_sections: yes
  html_document: default
geometry: margin=1.5in, top=0.5in, bottom=0.5in
---

Each student should provide a Rmd file with _two_ to _four_ plots, with text describing the semantics of the data, the question, how they have answered the question, and an explanation for each figure, showing how that particular figure helps the answering of the initial question.  Fork the LPS repository in GitHub, push your Rmd solution there. Send us, by e-mail, the link for your GIT repository, indicating the PATH to the Rmd file. Check the LPS website for the deadline.

# Introduction

The City of Porto Alegre, under the transparency law, has provided a data set with all the urban accidents (within the city limits) since 2000. The data set, including a description of each column in the PDF file format, is available in the following website:

http://www.datapoa.com.br/dataset/acidentes-de-transito

# Goal

For a given year (defined by the LPS coordination for each student enrolled in the cursus), the goal is to answer one of the following questions. The solution must use the data import and manipulation verbs of the R programming language and the tidyverse metapackage (readr, tidyr, dplyr) using Literate Programming.

# Questions

1. What is the time of the day with most accidents?
2. How many vehicles are involved in the accidents?
3. What types of accidents are more common?
4. Is the number of deaths increasing or decreasing?
5. Is there a street of the city with more accidents than others?
6. Do holidays impact in the number of accidents?

# Download the data

Supposing you have the URL for the CSV file, you can read the data using the code below. You can also download it manually and commit it to your repository to avoid an internet connection every time you knit this file. If the URL changes, the second solution might even make your analysis be more portable in time.

```{r}
library(readr)
data <- read_delim("/home/gauss/lllemos/lps/tasks/acidentes-2006.csv", delim=";")
data
```


Load the necessary packages:

```{r}
library(dplyr);
library(magrittr);
```

# Answer

The data used to develop the work is the urban accidents of 2006 from Porto Alegre. The dataset has specific information about the accidents.

The question choseen to be answered is "Q2. How many vehicles are involved in the accidents?"


First, it's selected the variables related to vehicles and the rows of these variables represent the amount of vehicles involved in the accidents.


```{r}
data %>% select(AUTO, TAXI, LOTACAO, ONIBUS_URB, ONIBUS_INT, CAMINHAO, MOTO, CARROCA, BICICLETA, OUTRO)

```


Second, the following plot shows during the year of 2006 how many vehicles of each type are involved in accidents.  We can notice that automobiles (variable 'AUTO') are the ones that are most involved in accidents. The x axis represents the classes of vehicles and the y axis the amount of these vehicles involved in accidents during the whole year.


```{r}
library(ggplot2);
library(reshape2);

new_data <- data %>% summarize(auto=sum(AUTO),  moto=sum(MOTO), taxi=sum(TAXI), lotacao=sum(LOTACAO), onibusur=sum(ONIBUS_URB), onibusint=sum(ONIBUS_INT), caminhao = sum(CAMINHAO), carroca=sum(CARROCA), bicicleta = sum(BICICLETA),outro=sum(OUTRO))

new_data <- melt(new_data[,c('auto', 'moto', 'taxi', 'lotacao', 'onibusur', 'onibusint', 'caminhao', 'carroca', 'bicicleta', 'outro')])


ggplot(new_data, aes(x=variable, y=value)) + geom_bar(stat = "identity") + ylim(0,NA) + theme_bw();

  
```



Third, it's used the verb mutate to create a new column(variable) with the sum of all types of vehicles involved in the accidents. After that, it's created a bar plot to see which interval has the highest incidence of accidents. Also, the mean and standart deviation are showed.

In the plot the x axis represents the new variable of how many vehicles are involved in each accidents and the y axis represents the amount of vehicles involved.

```{r}
library(ggplot2);

data %>%  mutate(sum_veh = AUTO+MOTO+LOTACAO+ONIBUS_URB, ONIBUS_INT, CAMINHAO, MOTO, CARROCA, BICICLETA, OUTRO) %>% summarize(mean_sum_veh = mean(sum_veh), std = sd(sum_veh))
```

```{r}
library(ggplot2);
data %>%  mutate(sum_veh = AUTO+MOTO+LOTACAO+ONIBUS_URB, ONIBUS_INT, CAMINHAO, MOTO, CARROCA, BICICLETA, OUTRO) %>% ggplot(aes(sum_veh)) + geom_bar() + ylim(0,NA) + theme_bw();

```





















